---
title: Using integrated multispecies occupancy models to map co-occurrence between
  bottlenose dolphins and fisheries in the Gulf of Lion, French Mediterranean Sea.
author: "Valentin Lauret, Hélène Labach, Léa David, Matthieu Authier, Olivier Gimenez"
subtitle: R codes
output:
  pdf_document: default
  word_document: default
  bookdown::pdf_document2:
    extra_dependencies: float
    highlight: tango
geometry: margin=0.5cm
toc: no
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval = FALSE, 
                      echo = TRUE,
                      results = 'hide', 
                      dpi = 300, 
                      cache = TRUE)
```

# Load required packages

```{r}
library(tidyverse)
library(sf)
library(nimble)
```

# Load data

```{r ,  message=FALSE, warning=FALSE}
load("MSOdata.rdata")
```

# MSO

First load the model 

## BUGS model

```{r eval = FALSE, include=TRUE }
MSO <- nimbleCode({

  ## state process
  for(j in 1:nsite){
    z[j] ~ dcat(psi[j, 1:4])
  }
  # occupancy probabilities

  psi[1:nsite, 1] <- 1 / (1 + sum(prop[1:nsite, 1:3])) # unoccupied
  psi[1:nsite, 2] <- prop[1:nsite, 1] / (1 + sum(prop[1:nsite, 1:3])) # occupied by species A and not B
  psi[1:nsite, 3] <- prop[1:nsite, 2] / (1 + sum(prop[1:nsite, 1:3])) # occupied by species B and not A
  psi[1:nsite, 4] <- prop[1:nsite, 3] / (1 + sum(prop[1:nsite, 1:3])) # occupied by both species A and B

  ## observation process
  for(j in 1:nsite) {
    for(k in 1:nyear) {
      y[j, k] ~ dcat(obs[j, k, 1:4, z[j]])
    }
  }

  # detection matrix with obs for observations and state = true states
  # obs take values:
  ## Let's have a bit of explanation, we have species A, and B. We have monitoring through SAMM only
  ## Then, we have 4 different situations
  # 1 for none species detected
  # 2 for species A
  # 3 for species B
  # 4 for both species detected

  # given state = unoccupied,
  for(j in 1:nsite) {
    for(k in 1:nyear) {
      # state 1 = no species use the site
      obs[j, k, 1,  1] <- 1 # prob obs = 1
      obs[j, k, 2,  1] <- 0 # prob obs = 2
      obs[j, k, 3,  1] <- 0 # prob obs = 3
      obs[j, k, 4,  1] <- 0 # prob obs = 4

      # given state 2 = occupied by species A and not B,
      obs[j, k, 1,  2] <- 1 - pA[j,k] # prob obs = 1
      obs[j, k, 2,  2] <- pA[j,k] # prob obs = 2
      obs[j, k, 3,  2] <- 0 # prob obs = 3
      obs[j, k, 4,  2] <- 0 # prob obs = 4

      # given state 3 = occupied by species B and not A,
      obs[j, k, 1,  3] <- 1 - pB[j,k] # prob obs = 1
      obs[j, k, 2,  3] <- 0 # prob obs = 2
      obs[j, k, 3,  3] <- pB[j,k] # prob obs = 3
      obs[j, k, 4,  3] <- 0 # prob obs = 4

      # given state 4 = occupied by both species B and A,
      obs[j, k, 1,  4] <- (1 - pA[j,k]) * (1 - pB[j,k]) # prob obs = 1
      obs[j, k, 2,  4] <-  pA[j,k] * (1 - pB[j,k]) # prob obs = 2
      obs[j, k, 3,  4] <-  pB[j,k] * (1 - pA[j,k]) # prob obs = 3
      obs[j, k, 4,  4] <-  pA[j,k] * pB[j,k]   # prob obs = 4
    }
  }
  ## priors for...
  # occupancy probabilities

  for(j in 1:nsite) {
    log(prop[j, 1]) <- theta1[j]
    log(prop[j, 2]) <- theta2[j]
    log(prop[j, 3]) <- theta3[j]
  }

  theta1[1:nsite] <- X[1:nsite,1:34] %*% b1[1:34] ## linear predictor
  theta2[1:nsite] <- X[1:nsite,1:34] %*% b2[1:34] ## linear predictor
  theta3[1:nsite] <- X[1:nsite,1:34] %*% b3[1:34] ## linear predictor


  b1[1] ~ dnorm(0,0.01)
  b2[1] ~ dnorm(0,0.01)
  b3[1] ~ dnorm(0,0.01)

  ## prior for s(bathy)
  b1[2] ~ dnorm(0,0.01)
  b2[2] ~ dnorm(0,0.01)
  b3[2] ~ dnorm(0,0.01)

  ## prior for s(coordx,coordy)
  K21[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 1]  + S1[1:32,33:64] * lambda[2, 1]
  K22[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 2]  + S1[1:32,33:64] * lambda[2, 2]
  K23[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 3]  + S1[1:32,33:64] * lambda[2, 3]
  b1[3:34] ~ dmnorm(zero[1:32], K21[1:32,1:32])
  b2[3:34] ~ dmnorm(zero[1:32], K22[1:32,1:32])
  b3[3:34] ~ dmnorm(zero[1:32], K23[1:32,1:32])

  ## smoothing parameter priors
  for (i in 1:2) {
    for (kk in 1:3){
      lambda[i, kk] ~ dgamma(.05,.005)
      rho[i, kk] <- log(lambda[i, kk])
    }
  }
  # detection probabilities (pA function of pres/abs of B, as in Waddle et al 2010 page 1470)
  # VL: There are 2 detections probabilities now pA, pB

  for(j in 1:nsite) {
    B_present[j] <- equals(z[j],3) + equals(z[j],4) #psi[j, 3] + psi[j, 4]

    for(k in 1:nyear) {
      pA[j, k] <- (1/(1 + exp(-(beta[1] * B_present[j] + beta[2] * (1 - B_present[j]) + beta[3] * eff[j, k]))))*effind[j,k]

      pB[j, k] <- (1/(1 + exp(-(beta[4] + beta[5] * eff[j, k]))))*effind[j,k]


    }
  }
  for (i in 1:5){
    beta[i] ~ dnorm(0,1)
  }
})

```


Then, you choose to prepare either SAMM or GDEGeM dataset for the run, ie. run only one of the next paragraph.

### SAMM

Get the ingredients for GAMs using package `jagam` developed by Simon Wood and basically hacks what is built by the package `mgcv`.

```{r message=FALSE, warning=FALSE}
yy_dolphinS <- apply(y2S, 1, max, na.rm = TRUE)
yy_dolphinS[yy_dolphinS ==1] <- 0
yy_dolphinS[yy_dolphinS >1] <- 1

coordxS <- coordMaskS[,1]
coordyS <- coordMaskS[,2]
stbathyS <- mt$bathy.sc[!ind_effortS,1]
library(mgcv)
res <- jagam(yy_dolphinS ~ stbathyS + s(coordxS, coordyS, bs = "gp"), 
             family = "binomial",
             file = "psi.txt") # same structure for fishing boats and both together
#save(res, file = 'jagam.RData')
```

```{r eval = FALSE, include=FALSE}
# nimble
y2S[is.na(y2S)] <-  1

data <- list(y = y2S, 
             eff = seffS,
             effind = effindS,
             X = res$jags.data$X, # gam para
             S1 = res$jags.data$S1,
             #S2 = res$jags.data$S2,
             zero = res$jags.data$zero)

constants <- list(nsite = dim(y2S)[1], 
                  nyear = dim(y2S)[2])

# Initial values for z {1:4} have to be consistent with the detection {1:16}

zinit <- matrix(NA, nrow = dim(y2S)[1], ncol = dim(y2S)[2])
zinit <- rep(1, dim(y2S)[1])

y2S[is.na(y2S)] <- 1
for (i in 1:dim(y2S)[1]){
  d <- f <- b <- 0
  for(j in 1:dim(y2S)[2]){
    if (y2S[i,j] == 2) {d <- 1}
    if (y2S[i,j] == 3) {f <- 1} 
    if (y2S[i,j] == 4) {b <- 1}
  } #j
  if(d == 1 & f == 0 & b == 0) {zinit[i] <- 2}
  if(d == 0 & f == 1 & b == 0) {zinit[i] <- 3}
  if(b == 1) {zinit[i] <- 4}
  if(d == 1 & f ==1) {zinit[i] <- 4}
}#i

# table
table(zinit)

inits <- list(z = zinit, 
              beta = rnorm(5, 0, 1), 
              lambda = cbind(res$jags.ini$lambda, res$jags.ini$lambda, res$jags.ini$lambda),
              b1 = res$jags.ini$b,
              b2 = res$jags.ini$b,
              b3 = res$jags.ini$b)
```


### GDEGeM

Get the ingredients for GAMs using package `jagam` developed by Simon Wood and basically hacks what is built by the package `mgcv`.

```{r message=FALSE, warning=FALSE}
yy_dolphinG <- apply(y2G, 1, max, na.rm = TRUE)
yy_dolphinG[yy_dolphinG ==1] <- 0
yy_dolphinG[yy_dolphinG >1] <- 1

coordxG <- coordMaskG[,1]
coordyG <- coordMaskG[,2]
stbathyG <- mt$bathy.sc[!ind_effortG,1]
library(mgcv)
res <- jagam(yy_dolphinG ~ stbathyG + s(coordxG, coordyG, bs = "gp"), 
             family = "binomial",
             file = "psi.txt") # same structure for fishing boats and both together
#save(res, file = 'jagam.RData')
```

```{r eval = FALSE, include=FALSE}
# nimble
y2G[is.na(y2G)] <-  1

data <- list(y = y2G, 
             eff = seffG,
             effind = effindG,
             X = res$jags.data$X, # gam para
             S1 = res$jags.data$S1,
             #S2 = res$jags.data$S2,
             zero = res$jags.data$zero)

constants <- list(nsite = dim(y2G)[1], 
                  nyear = dim(y2G)[2])

# Initial values for z {1:4} have to be consistent with the detection {1:16}

zinit <- matrix(NA, nrow = dim(y2G)[1], ncol = dim(y2G)[2])
zinit <- rep(1, dim(y2G)[1])

y2G[is.na(y2G)] <- 1
for (i in 1:dim(y2G)[1]){
  d <- f <- b <- 0
  for(j in 1:dim(y2G)[2]){
    if (y2G[i,j] == 2) {d <- 1}
    if (y2G[i,j] == 3) {f <- 1} 
    if (y2G[i,j] == 4) {b <- 1}
  } #j
  if(d == 1 & f == 0 & b == 0) {zinit[i] <- 2}
  if(d == 0 & f == 1 & b == 0) {zinit[i] <- 3}
  if(b == 1) {zinit[i] <- 4}
  if(d == 1 & f ==1) {zinit[i] <- 4}
}#i

# table
table(zinit)

inits <- list(z = zinit, 
              beta = rnorm(5, 0, 1), 
              lambda = cbind(res$jags.ini$lambda, res$jags.ini$lambda, res$jags.ini$lambda),
              b1 = res$jags.ini$b,
              b2 = res$jags.ini$b,
              b3 = res$jags.ini$b)
```




## Build, compile and run model with NIMBLE

```{r eval = FALSE, include=TRUE}
Rmodel <- nimbleModel(IMSO, constants, data, inits)
Rmodel$initializeInfo()
Rmodel$calculate()

conf <- configureMCMC(Rmodel)
conf$printMonitors() 
conf$addMonitors("z")

conf$printSamplers(byType= TRUE)

# Build and compile MCMC
Rmcmc <- buildMCMC(conf)
Cmodel <- compileNimble(Rmodel)
Cmcmc <- compileNimble(Rmcmc, project = Cmodel)

t <- system.time(samples2 <- runMCMC(Cmcmc, niter = 100000, nburnin = 10000, nchains = 3,
                                     samplesAsCodaMCMC = TRUE))
```
