---
title: "MSO_integrated"
author: "Valentin Lauret"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', 
                      dpi = 300, 
                      cache = TRUE)
library(tidyverse)
theme_set(theme_light())
library(sf)
library(nimble)
```

# Format and visualise data 

Load grid and occupancy data.
```{r message=FALSE, warning=FALSE}
load("pays.rdata")
load("~/Google Drive/These/Work/Occupancy/sea.grid.rdata")
  
grid <- st_read("Grid/grid.shp") %>% st_transform(crs = st_crs(pays))
# focus Golfe du Lion
grid <-  grid %>% st_crop(xmin = 700000, xmax = 900000, ymin = 6140000, ymax = 6300000) 
pays <- pays %>% st_crop(st_bbox(grid))

sea <- st_transform(sea, crs = st_crs(grid))
mt <- st_crop(sea,grid)


# GDEGeM data
load('msoccu_gd.rdata')
multioccu_gd$dauphins <- st_transform(multioccu_gd$dauphins, st_crs(grid))
multioccu_gd$chalut <-   st_transform(multioccu_gd$chalut, st_crs(grid))
multioccu_gd$effort <-   st_transform(multioccu_gd$effort, st_crs(grid))
```

## GDEGeM
Visualise data.
```{r message=FALSE, warning=FALSE}
dfG <- bind_rows(multioccu_gd$dauphins %>% add_column(species = "dauphin"), 
          multioccu_gd$chalut %>% add_column(species = "chalutier")) %>% 
  st_crop(st_bbox(grid))

```

Build datasets. 
```{r message=FALSE, warning=FALSE}
# sampling effort
effortG <- multioccu_gd$effort %>% 
  st_crop(st_bbox(grid)) %>% 
  select(autumn:summer) %>%
  as_tibble() %>%
  select(-geometry) %>% 
  as.matrix()

# dolphin detections/non-detections

y_dolphinG <- dfG %>% 
  filter(species == "dauphin") %>%
  select(autumn:summer) %>%
  as_tibble() %>%
  select(-geometry) %>% 
  as.matrix()
y_dolphinG[effortG == 0] <- NA
## ind_dolphinG <- apply(y_dolphinG, 1, function(x) all(is.na(x)))
## y_dolphinG <- y_dolphinG[ !ind_dolphinG, ]

# fishing boats detections/non-detections
y_fishingG <- dfG %>% 
  filter(species == "chalutier") %>%
  select(autumn:summer) %>%
  as_tibble() %>%
  select(-geometry) %>% 
  as.matrix()
y_fishingG[effortG == 0] <- NA
## ind_fishingG <- apply(y_fishingG, 1, function(x) all(is.na(x)))
## y_fishingG <- y_fishingG[ !ind_fishingG, ]

# grid cells coordinates
coordG <- dfG %>% 
  filter(species == "dauphin") %>%
  select(autumn:summer) %>%
  st_centroid() %>%
  st_coordinates() %>%
  as_tibble() %>%
  mutate(easting = (X - mean(X)) / sd(X), 
         northing = (Y - mean(Y)) / sd(Y)) %>%
  select(easting, northing) %>%
  as.matrix()
## maskG <- apply(effortG == 0, 1, sum) == 4
## coordG <- coordG[!maskG,]
dim(coordG)

# means and standard deviations used to standardise the grid cells coordinates
tempG <- dfG %>% 
  filter(species == "dauphin") %>%
  st_centroid() %>%
  st_coordinates() %>%
  as.matrix() 
meanXG <- mean(tempG[,1])
sdXG <- sd(tempG[,1])
meanYG <- mean(tempG[,2])
sdYG <- sd(tempG[,2])
```

## Joining the dataset and formatting for Bayesian analyses

Format data.
```{r message=FALSE, warning=FALSE}
yG <- list(y_dolphinG, y_fishingG)

## ind_effortG <- apply(effortG, 1, sum)
cov_effortG <- effortG
cov_effortG[cov_effortG == 0] <- NA
st_effortG <- matrix(scale(cov_effortG), ncol = 4)
```

Let's format the data in a matrix with $N$ rows (sites) and $J$ columns (surveys) with in each cell a `1:16` for the observation (or event in the capture-recapture terminology) `none species detected`, `species A detected`, `species B detected`, `both species detected` and each of these event by either by SAMM only, by GDEGeM onyl, or by both monitoring programs.  

```{r message=FALSE, warning=FALSE}
str(yG)
N <- nrow(yG[[1]])
J <- ncol(yG[[1]])
 y_jagsG  <-  matrix(NA, nrow = N, ncol = J)


# GDEGeM
for (j in 1:N){ #nsites
  for (k in 1:J){ #nocc
    if (is.na(yG[[1]][j,k])) next # if cell j is not sampled at occasion k, then next
    if (yG[[1]][j,k] == 0 & yG[[2]][j,k] == 0) y_jagsG[j,k] <- 1
    if (yG[[1]][j,k] == 1 & yG[[2]][j,k] == 0) y_jagsG[j,k] <- 2
    if (yG[[1]][j,k] == 0 & yG[[2]][j,k] == 1) y_jagsG[j,k] <- 3
    if (yG[[1]][j,k] == 1 & yG[[2]][j,k] == 1) y_jagsG[j,k] <- 4
  }
}

dim(y_jagsG)
dim(st_effortG)
# Before joining, we do a trick to deal with NAs. We replace the NAs by 0 in where the couple site x occasion has been sampled by the other monitoring device



# apply mask to remove site without sampling effort from none of the monitoring program
 ind_effort <- apply(y_jagsG, 1, function(x) all(is.na(x)))
 y <- y_jagsG[ !ind_effort, ]
 seffG <- st_effortG[ !ind_effort, ]
 effindG <- seffG
 effindG[is.na(effindG)] <- 0
 effindG[effindG !=0] <- 1
 
 
 seffG[is.na(seffG)] <- 0

# same mask with coordG,coordS
 coordMask <- coordG[ !ind_effort, ]
mt <-  mt[!ind_effort,]
```

# Simplest MSO

## BUGS model 

### JAGS

Let's introduce the GAM

```{r}
yy_dolphin <- apply(y, 1, max, na.rm = TRUE)
yy_dolphin[yy_dolphin ==1] <- 0
yy_dolphin[yy_dolphin >1] <- 1

coordx <- coordMask[,1]
coordy <- coordMask[,2]
library(mgcv)
res <- jagam(yy_dolphin ~ s(coordx, coordy, bs = "gp"), 
            family = "binomial",
            file = "psi.txt") # same structure for fishing boats and both together
#save(res, file = 'jagam.RData')
```



I did not display JAGS process although I made it work in the .Rmd file.  

```{r eval = FALSE, include=FALSE}
MSO <- function(){ 
  
  ## state process
  for(j in 1:nsite) {

    z[j] ~ dcat(psi[j, 1:4])
  }
  # occupancy probabilities
  for (i in 1:nsite){
  psi[i, 1] <- 1 / (1 + sum(prop[i, 1:3])) # unoccupied
  psi[i, 2] <- prop[i, 1] / (1 + sum(prop[i, 1:3])) # occupied by species A and not B
  psi[i, 3] <- prop[i, 2] / (1 + sum(prop[i, 1:3])) # occupied by species B and not A
  psi[i, 4] <- prop[i, 3] / (1 + sum(prop[i, 1:3])) # occupied by both species A and B
  }

  ## observation process
  for(j in 1:nsite) {
    for(k in 1:nyear) {
      y[j, k] ~ dcat(obs[j, k, 1:4, z[j]])
    }
  }
  
  # detection matrix with obs for observations and state = true states
  # obs take values:
  # 1 for none species detected
  # 2 for species 1 detected
  # 3 for species 2 detected
  # 4 for both species detected
  # given state = unoccupied,
  for(j in 1:nsite) {
    for(k in 1:nyear) {
      obs[j, k, 1, 1] <- 1 # prob obs = 1 
      obs[j, k, 2, 1] <- 0 # prob obs = 2
      obs[j, k, 3, 1] <- 0 # prob obs = 3
      obs[j, k, 4, 1] <- 0 # prob obs = 4
      # given state = occupied by species A and not B,
      obs[j, k, 1, 2] <- 1 - pA[j, k] # prob obs = 1
      obs[j, k, 2, 2] <- pA[j, k] # prob obs = 2
      obs[j, k, 3, 2] <- 0 # prob obs = 3
      obs[j, k, 4, 2] <- 0 # prob obs = 4
      # given state = occupied by species B and not A,
      obs[j, k, 1, 3] <- 1 - pB[j, k] # prob obs = 1
      obs[j, k, 2, 3] <- 0 # prob obs = 2
      obs[j, k, 3, 3] <- pB[j, k] # prob obs = 3
      obs[j, k, 4, 3] <- 0 # prob obs = 4
      # given state = occupied by both species A and B,
      obs[j, k, 1, 4] <- (1 - pA[j, k]) * (1 - pB[j, k]) # prob obs = 1
      obs[j, k, 2, 4] <- pA[j, k] * (1 - pB[j, k]) # prob obs = 2
      obs[j, k, 3, 4] <- (1 - pA[j, k]) * pB[j, k] # prob obs = 3
      obs[j, k, 4, 4] <- pA[j, k] * pB[j, k] # prob obs = 4
    }
  }  
  ## priors for...
  # occupancy probabilities

 for(j in 1:nsite) {
    log(prop[j, 1]) <- theta1[j] 
    log(prop[j, 2]) <- theta2[j] 
    log(prop[j, 3]) <- theta3[j]
  }
  
  theta1[1:nsite] <- X[1:nsite,1:33] %*% b1[1:33] ## linear predictor
  theta2[1:nsite] <- X[1:nsite,1:33] %*% b2[1:33] ## linear predictor
  theta3[1:nsite] <- X[1:nsite,1:33] %*% b3[1:33] ## linear predictor

    b1[1] ~ dnorm(0,0.01) 
    b2[1] ~ dnorm(0,0.01) 
    b3[1] ~ dnorm(0,0.01) 

  ## prior for s(coordx,coordy)... 
  K11[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 1]  + S1[1:32,33:64] * lambda[2, 1]
  K12[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 2]  + S1[1:32,33:64] * lambda[2, 2]
  K13[1:32,1:32] <- S1[1:32,1:32] * lambda[1, 3]  + S1[1:32,33:64] * lambda[2, 3]
  b1[2:33] ~ dmnorm(zero[2:33], K11[1:32,1:32]) 
  b2[2:33] ~ dmnorm(zero[2:33], K12[1:32,1:32]) 
  b3[2:33] ~ dmnorm(zero[2:33], K13[1:32,1:32]) 
  ## smoothing parameter priors CHECK...
  for (i in 1:2) {
    for (kk in 1:3){
      lambda[i, kk] ~ dgamma(.05,.005)
      rho[i, kk] <- log(lambda[i, kk])
    }
  } 
    # detection probabilities (pA function of pres/abs of B, as in Waddle et al 2010 page 1470)
  for(j in 1:nsite) {
    
    B_present[j] <- equals(z[j], 3) + equals(z[j], 4)
    
    for(k in 1:nyear) {
      logit(pA[j, k]) <- ifelse(effind[j,k], beta[1] + beta[2]*B_present[j] + beta[3] * eff[j, k] ,0)
      logit(pB[j, k]) <- ifelse(effind[j,k], beta[4] + beta[5] * eff[j, k],0)
    }
  }
  for (i in 1:5){
    beta[i] ~ dnorm(0,1)
  }
}
```

Initial values.
```{r eval = FALSE , include=FALSE}
# Initial values for z {1:4} have to be consistent with the detection {1:16}
# To remain simple with put z = 4 as initial value if smth is detected
# Serious inits
zinit <- matrix(NA, nrow = dim(y)[1], ncol = dim(y)[2])
zinit <- rep(1, dim(y)[1])

for (i in 1:dim(y)[1]){
  d <- f <- b <- 0
  for(j in 1:dim(y)[2]){
    if (y[i,j] == 2) {d <- 1}
    if (y[i,j] == 3) {f <- 1} 
    if (y[i,j] == 4) {b <- 1}
    } #j
  if(d == 1 & f == 0 & b == 0) {zinit[i] <- 2}
  if(d == 0 & f == 1 & b == 0) {zinit[i] <- 3}
  if(b == 1) {zinit[i] <- 4}
  if(d == 1 & f ==1) {zinit[i] <- 4}
}#i


y[is.na(y)] <- 1

seffG[is.na(seffG)] <- 0
data <- list(y = y, 
             nsite = dim(y)[1], 
             nyear = dim(y)[2],
             eff = seffG,
             effind = effindG,
             X = res$jags.data$X, # gam para
             S1 = res$jags.data$S1,
             zero = res$jags.data$zero)

inits <- function() {list(z = zinit, 
                          beta = rnorm(5, 0, 1), 
                          lambda = cbind(res$jags.ini$lambda, res$jags.ini$lambda, res$jags.ini$lambda),
              b1 = res$jags.ini$b,
              b2 = res$jags.ini$b,
              b3 = res$jags.ini$b)}
```

Parameters to be monitored.
```{r eval = FALSE, include=FALSE}
params <- c("prop","beta","b1","b2","b3")
```

MCMC settings
```{r eval = FALSE, include=FALSE}
ni <- 4000
nb <- 250
nc <- 1
```

Run `Jags` from `R`.
```{r eval = FALSE, include=FALSE}
library(R2jags)
ptm <- proc.time()
out2 <- jags(data = data, 
            inits = inits, 
            parameters.to.save = params, 
            model.file = MSO, 
            n.chains = nc, 
            n.iter = ni, 
            n.burnin = nb)
x2 <- proc.time() -  ptm
x2 # 15 minutes

mcmcplots::denplot(out2$BUGSoutput$sims.list$beta)
```

## Deal with output

```{r}
beta1 <- c(out2$BUGSoutput$sims.array[,,'beta[1]']) 
beta2 <- c(out2$BUGSoutput$sims.array[,,'beta[2]']) 
beta3 <- c(out2$BUGSoutput$sims.array[,,'beta[3]'])
beta4 <- c(out2$BUGSoutput$sims.array[,,'beta[4]'])
beta5 <- c(out2$BUGSoutput$sims.array[,,'beta[5]'])
```

```{r}
grid_p <- seq(range(seffG)[1], range(seffG)[2], length = 100) 

logit_p12 <- median(beta1) + median(beta2) + median(beta3) * grid_p
logit_p12bar <- median(beta1) + median(beta3) * grid_p
logit_p2 <- median(beta4) + median(beta5) * grid_p

ggplot()+ 
  geom_line(aes(x = grid_p, y = plogis(logit_p12), color = "dauphin cond présence chalutier"),  lwd = 1) + 
  geom_line(aes(x = grid_p, y = plogis(logit_p12bar), color = "dauphin cond absence chalutier"), lwd = 1) + 
  geom_line(aes(x = grid_p, y = plogis(logit_p2), color = "chalutier"), lwd = 1) +
  scale_color_manual(values = c("dauphin cond présence chalutier" = "#fb8b24", "dauphin cond absence chalutier" = "#0F4C5C" , "chalutier" = "#9a031e")) + 
  xlab("effort (km parcouru)") + ylab("probabilité estimée") + labs(title = "GDEGeM : probabilité de détection pour un...", color = "Legend")
```

```{r}
grid_b <- mt$bathy.sc[,1]

prop1 <- exp(median(alpha1) + median(alpha2) * grid_b)
prop2 <- exp(median(alpha3) + median(alpha4) * grid_b)
prop3 <- exp(median(alpha5) + median(alpha6) * grid_b) 

psi1 <- plogis(prop1) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi2 <- plogis(prop2) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi3 <- plogis(prop3) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi0 <- 1 - (psi1 + psi2 + psi3)  
```

First, get the whole grid.

```{r echo=TRUE}
grid_coord <- grid %>%
  st_centroid() %>%
  st_coordinates() %>%
  as_tibble() %>%
  mutate(easting = (X - meanXG)/sdXG,
         northing = (Y - meanYG)/sdYG)

```

Second, get linear predictor.

```{r echo=TRUE}
sm <- smoothCon(s(coordx, coordy, bs = "gp"), 
                          data = data.frame(coordx = grid_coord$easting, 
                                            coordy = grid_coord$northing), 
                          absorb.cons = TRUE) 

Xp <- PredictMat(sm[[1]], data.frame(coordx = grid_coord$easting, 
                                     coordy = grid_coord$northing))
Xp <- cbind(1, Xp)

b1 <- out2$BUGSoutput$sims.list$b1
b2 <- out2$BUGSoutput$sims.list$b2
b3 <- out2$BUGSoutput$sims.list$b3


#dim(Xp)
#dim(b1)

mu1 <- matrix(NA, nrow = nrow(Xp), ncol = nrow(b1))
mu2 <- matrix(NA, nrow = nrow(Xp), ncol = nrow(b2))
mu3 <- matrix(NA, nrow = nrow(Xp), ncol = nrow(b3))
for (i in 1:nrow(b1)){
  mu1[1:nrow(Xp), i] <- Xp %*% b1[i,]
  mu2[1:nrow(Xp), i] <- Xp %*% b2[i,]
  mu3[1:nrow(Xp), i] <- Xp %*% b3[i,]
}

prop1 <- apply(exp(mu1), 1, mean)
prop2 <- apply(exp(mu2), 1, mean)
prop3 <- apply(exp(mu3), 1, mean)

psi1 <- plogis(prop1) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi2 <- plogis(prop2) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi3 <- plogis(prop3) / (1 + plogis(prop1) + plogis(prop2) + plogis(prop3))
psi0 <- 1 - (psi1 + psi2 + psi3)
```


```{r}
ggplot()+ 
  geom_sf(data = grid,aes(fill = psi3),  lwd = 0.1) + 
  geom_sf(data = pays) 

# pr dauphin
tur <- ggplot()+ 
  geom_sf(data = grid,aes(fill = (psi1 + psi3)),  lwd = 0.1) + 
  geom_sf(data = pays) +scale_fill_viridis_c()

# pr chalut
cha <- ggplot()+ 
  geom_sf(data = grid,aes(fill = (psi2 + psi3)),  lwd = 0.1) + 
  geom_sf(data = pays)  + 
  scale_fill_viridis_c()

(tur/cha)
```

`


